Intrusion Detection with Genetic Algorithms and Fuzzy Logic

### 2013-10-27 paper.tex
Wrote some things in the genetic algorithm section, also wrote about true positives and true negatives.

\subsection{Fuzzy Logic}
Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual. A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set. An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual. To determine what is considered high, the values of the packets are divided into fuzzy sets. In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C). 
\begin{figure}
\centering
\psfig{file=triangleFigure.JPG}
\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)}
\label{fig:triangleFigure}
\end{figure}
The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region. For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees. If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}

\subsection{Genetic Algorithm}

@@

\subsection{False Positives, False Negatives, True Positives, True Negatives}
\label{sec:PosNeg}
In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}. The given algorithm is then trained on the training set to look for patterns. These patterns are then verified using the test set is used.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.

\emph{False positives} and \emph{False negatives} are two concepts that are used to evaluate the accuracy of an intrusion detection system. A false positive happens when an intrusion detection system incorrectly identifies normal activity as being an attack. A false negative happens when an intrusion detection system fails to identify harmful activity.~\cite{Liao201316}

\mycomment{Write more here about True Pos, True Neg, Detection Rate}

@@

\mycomment{Fix tables}

>>>>

\subsection{Fuzzy Logic}
Attacks on systems do not always have a fixed patter, so fuzzy logic is used to detect patterns that have a behavior that is between normal and unusual. A fuzzy logic rule has the following format: If <\emph{condition}> then <\emph{consequence}>, where \emph{condition} is a fuzzy variable, and \emph{consequence} is a fuzzy set. An example of this is: if the number of packets with the same destination address is high, then the pattern is unusual. To determine what is considered high, the values of the packets are divided into fuzzy sets. In Figure~\ref{fig:triangleFigure} there are three sets: low (region A), medium (region B), and high (region C). 
\begin{figure}
\centering
\psfig{file=triangleFigure.JPG}
\caption{Fuzzy shape with three sets: low (A), medium (B), and high (C)~\cite{DBLP:journals/corr/abs-1204-6416}}
\label{fig:triangleFigure}
\end{figure}
The x-axis are the values in the fuzzy set (number of packets), and the y-axis is the membership function, which determines the region. For example, if the number of packets with the same destination address is 15 then this will be considered as low for a degree of 0.4, but will be considered medium or high for other degrees. If the number of packets is determined to be high, then the connection will be aborted.~\cite{DBLP:journals/corr/abs-1204-6416}

\subsection{Genetic Algorithm}
Genetic algorithms are a search technique used to find solutions to problems \emph{Mutation}, \emph{selection}, and \emph{crossover} are used to evolve and improve rules.

Mutation is where random bits in a solution are changed. Selection is where solutions that have a better \emph{fitness} are chosen over the other solutions. The fitness function determines the quality of a particular solution. Crossover is where two solutions swap one of their characteristics with the other to form two new solutions.

The solution to a problem is represented as a chromosome. First, a randomly generated population of chromosomes is created. According to the characteristics of the problem, the positions of each chromosome are encoded as bits, characters, or numbers. Then mutation, selection, and crossover are applied to each generation and eventually the best solution is found.~\cite{DBLP:journals/corr/abs-1204-6416}

@@

\subsection{False Positives, False Negatives, True Positives, True Negatives}
\label{sec:PosNeg}
In a machine learning experiment, a common technique is to divide the data set into two subsets, a \emph{training set} and a \emph{testing set}. The given algorithm is then trained on the training set to look for patterns. These patterns are then verified using the test set.~\cite{bc1_ecindm} Four different measures are then used to determine the accuracy of the algorithm in question.

A \emph{false positive} (FP) happens when an intrusion detection system incorrectly identifies normal activity as being an attack. A \emph{false negative} (FN) happens when an intrusion detection system fails to identify harmful activity. A \emph{true positive} (TP) happens when an intrusion detection system correctly identifies activities to be attacks. A \emph{true negative} (TN) happens when an intrusion detection system correctly identifies activities to be normal.

The \emph{detection rate} (DR) of an intrusion detection system is the number of true positives divided by the total number of intrusions that happen.~\cite{ids}

@@
